### 2018-01-26 dpCTL.tex

The logic $\dpCTL$ is designed to specify differentially private
mechanisms. We introduce the differentially private path quantifer
$\dpriv{\epsilon}{\delta}$ in $\dpCTL$. For any path formula
$\phi$, a state $s$ in a Markov chain $K$ satisfies
$\dpriv{\epsilon}{\delta} \phi$ if the probability of having $\phi$-paths 
from $s$ is close to the probabilities of 
having $\phi$-paths from its neighboring states. 
We give semantics for both Markov chains and Markov decision processes.