### 2018-01-27 dpCTL.tex
structures and some examples

The logic $\dpCTLstar$ is designed to specify differentially private
mechanisms. We introduce the differentially private path quantifer
$\dpriv{\epsilon}{\delta}$ in $\dpCTLstar$. For any path formula
$\phi$, a state $s$ in a Markov chain $K$ satisfies
$\dpriv{\epsilon}{\delta} \phi$ if the probability of having $\phi$-paths
from $s$ is close to the probabilities of
having $\phi$-paths from its neighboring states.
%We give semantics for both Markov chains and Markov decision processes.