### 2018-01-28 applications.tex
add frequency estimator

Consider the MDP in Figure~\ref{figure:simple-mdp-mdp}. Suppose the
states $+$ and $-$ are neighbors. Let us compute $\{ s : M,
\neighbor{S}, s \models \dpriv{3}{0} (\F Y) \}$. First, the self
product $M^2$ is depicted in Figure~\ref{figure:simple-mdp-product}
(the states $++$ and $--$ and associated transitions are ignored
for clarity). Applying
Theorem~\ref{theorem:eventuallyC} with $\epsilon = 4$ and $\delta = 0$, 
consider the following linear programming problem:
\[
\begin{array}{rcrcrcrcrcl}
  \multicolumn{11}{c}{
  \min x_{++} + x_{+-} + x_{-+} + x_{--} + x_{YY} + x_{YN} + x_{NY} + x_{NN}
  }\\
  \multicolumn{9}{r}{ x_{YN} } & = & 1 \\
  \multicolumn{9}{r}{ x_{YY} } & = & -3 \\
  \multicolumn{9}{r}{ x_{NY} } & = & -4 \\
  \multicolumn{9}{r}{ x_{NN} } & = & 0 \\
  x_{++} &-&   .5625 x_{YY} &-& .1875 x_{YN} &-& .1875 x_{NY} &-&
               .0625 x_{NN} & \geq & 0\\
  x_{++} &-&   .64   x_{YY} &-& .16   x_{YN} &-& .16   x_{NY} &-&
               .04   x_{NN} & \geq & 0\\
  x_{+-} &-&   .1875 x_{YY} &-& .5625 x_{YN} &-& .0625 x_{NY} &-& 
               .1875 x_{NN} & \geq & 0\\
  x_{+-} &-&   .16   x_{YY} &-& .64   x_{YN} &-& .04   x_{NY} &-&
               .16   x_{NN} & \geq & 0\\
  x_{-+} &-&   .1875 x_{YY} &-& .0625 x_{YN} &-& .5625 x_{NY} &-& 
               .1875 x_{NN} & \geq & 0\\
  x_{-+} &-&   .16   x_{YY} &-& .04   x_{YN} &-& .64   x_{NY} &-&
               .16   x_{NN} & \geq & 0\\
  x_{--} &-&   .0625 x_{YY} &-& .1875 x_{YN} &-& .1875 x_{NY} &-&
               .5625 x_{NN} & \geq & 0\\
  x_{--} &-&   .04   x_{YY} &-& .16   x_{YN} &-& .16   x_{NY} &-&
               .64   x_{NN} & \geq & 0
\end{array}
\]
The optimum is attained at 
$(x_{++}, x_{+-}, x_{-+}, x_{--}) = (-2.25, 0, -2.75, -0.6)$. By
Theorem~\ref{theorem:eventuallyC}, we have
\[
\begin{array}{l}
\max_{\scheduler{S}}
\myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
4 \myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
= -2.25 \leq 0\\
\max_{\scheduler{S}}
\myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
4 \myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
= 0 \leq 0 \\
\max_{\scheduler{S}}
\myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
4 \myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
= -2.75 \leq 0\\
\max_{\scheduler{S}}
\myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
4 \myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
= -0.6 \leq 0
\end{array}
\]
Therefore $+, - \in \{ s : M, \neighbor{M}, s \models \dpriv{3}{0}
(\F Y) \}$. This is expected. Recall that low- and high-accuracy
mechanisms are $(3, 0)$- and $(4, 0)$-differentially private
respectively. Regardless of request types, the interactive mechanism
in Figure~\ref{figure:simple-mdp-mdp} can only achieve $(4,
0)$-differential privacy at best.

Streaming processing gives another motivating example for Markov
decision processes. In stream processing, input data are retrieved and
analyzed continuously. Such continual observation is commonly found
in social media, service providers, or even electronic surveillance
systems. To protect privacy, privacy-respecting mechanisms must be
formulated and designed. Classical differential privacy however is
insufficient in streaming processing. Note that observations are of
indefinite lengths in such applications. All differentially private
mechanism necessarily stay in their internal states indefinitely
before outcomes are produced. Although differential privacy demands
privacy-respecting outcomes, internal states may reveal private
information in such mechanisms. When differentially private mechanisms
are in their internal states, privacy is still lost if an intrusion
occurs.

In pan privacy, data analysis mechanisms are required to respect
privacy even in their internal states~\cite{DNPR:10:DPCO,DNPRY:10:PPSA}. 
The idea is to show reachability probabilities are similar on
neighboring observations. In the following, we model a discrete
variant of the mechanisms by Markov decision processes. We will also
specify pan privacy in our model using $\dpCTL$.

Consider monitoring events of the types $a$ and $b$ continuously.
A sequence of event types is said to be an observation. Two
obsevations are {neighbors} if one can be obtained from the other 
by replacing event types any number of times, that is, if they are of
the same length. \bw{Later, we will consider a different definition of
neighborhood by restricting the number of replacements.} 
We would like to inquire the number of event types in
observations. For instance, the query result result is $1$ 
for observations $a \cdots a$ and $b \cdots b$ of arbitrary positive
lengths; and it is $2$ for other non-empty observations.

Based on the density estimation algorithm in \cite{DNPRY:10:PPSA}, we
consider the following mechanism for event type streams. Our mechanism
keeps a flag for each event type.

\begin{algorithm}
  \begin{algorithmic}[1]
  \Procedure{FrequencyEstimator}{}
    \State{$f_a \leftarrow 0, 1$ with probability 
      $\frac{1}{2}, \frac{1}{2}$ respectively}
    \Comment{Initialization}
    \State{$f_b \leftarrow 0, 1$ with probability
      $\frac{1}{2}, \frac{1}{2}$ respectively}
    \Match{current event type $e$}
    \Comment{Processing}
        \Case{$a$}
          \State{$f_a \leftarrow 0, 1$ with probability
            $\frac{3}{8}, \frac{5}{8}$ respectively}
        \EndCase
        \Case{$b$}
          \State{$f_b \leftarrow 0, 1$ with probability
            $\frac{3}{8}, \frac{5}{8}$ respectively}
        \EndCase
    \EndMatch
    \Match{$f_a + f_b$}
    \Comment{Output}
      \Case{0}
        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
          $\frac{2}{3}, \frac{1}{6}, \frac{1}{6}$ respectively}
      \EndCase
      \Case{1}
        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
          $\frac{1}{3}, \frac{1}{3}, \frac{1}{3}$ respectively}
      \EndCase
      \Case{2}
        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
          $\frac{1}{6}, \frac{1}{6}, \frac{2}{3}$ respectively}
      \EndCase
    \EndMatch
    \State{\Return $\mathit{out}$}
  \EndProcedure
  \end{algorithmic}
  \caption{FrequencyEstimator}
  \label{algorithm:frequency-estimator}
\end{algorithm}

For each event, a bit denotes whether it has occurred or not. To model
adjacent inputs, we use four actions symbols to denote events and its
replacement. For instance, the action $a[a]$ means the event $a$ is
observed without replacement; $b[a]$ means the event $b$ is observed
but replaced with $a$ in an adjacent input.

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
      \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (I) at  ( 0,  0) { $I$ };
      \node[node] (00) at (-1.5,  1.5) { $00$ };
      \node[node] (01) at (-1.5, -1.5) { $01$ };
      \node[node] (10) at ( 1.5,  1.5) { $10$ };
      \node[node] (11) at ( 1.5, -1.5) { $11$ };

      \path
      (I) edge node [below=2] { $1/4$ } (00)
      (I) edge node [above=2] { $1/4$ } (01)
      (I) edge node [below=2] { $1/4$ } (10)
      (I) edge node [above=2] { $1/4$ } (11)

      (00) edge [loop above] node { $a[\uscore]$, .375 } (00)
      (00) edge [bend left=10] node { $a[\uscore]$, .625 } (10)
      (00) edge [loop left] node [above=1,rotate=90] { $b[\uscore]$, .375 } (00)
      (00) edge [bend left=10] node [left=3,above=1,rotate=270] { $b[\uscore]$, .625 } (01)

      (10) edge [loop above] node { $a[\uscore]$, .625 } (00)
      (10) edge [bend left=10] node { $a[\uscore]$, .375 } (00)
      (10) edge [loop right] node [above=1,rotate=270] { $b[\uscore]$, .375 } (10)
      (10) edge [bend left=10] node [above=1,rotate=270] { $b[\uscore]$, .625 } (11)

      (11) edge [loop below] node { $a[\uscore]$, .625 } (11)
      (11) edge [bend left=10] node { $a[\uscore]$, .375 } (01)
      (11) edge [loop right] node [above=1,rotate=270] { $b[\uscore]$, .625 } (11)
      (11) edge [bend left=10] node [left=-3,above=1,rotate=90] { $b[\uscore]$, .375 } (10)

      (01) edge [loop below] node { $a[\uscore]$, .375 } (01)
      (01) edge [bend left=10] node { $a[\uscore]$, .675 } (11)
      (01) edge [loop left] node [above=1,rotate=90] { $b[\uscore]$, .625 } (01)
      (01) edge [bend left=10] node [left=-3,above=1,rotate=90] { $b[\uscore]$, .375 } (00)

      ;

      \node[node] (I') at  ( 5.5,    0) { $\underline{I}$ };
      \node[node] (00') at (   4,  1.5) { $\underline{00}$ };
      \node[node] (01') at (   4, -1.5) { $\underline{01}$ };
      \node[node] (10') at (   7,  1.5) { $\underline{10}$ };
      \node[node] (11') at (   7, -1.5) { $\underline{11}$ };

      \path
      (I') edge node [below=2] { $1/4$ } (00')
      (I') edge node [above=2] { $1/4$ } (01')
      (I') edge node [below=2] { $1/4$ } (10')
      (I') edge node [above=2] { $1/4$ } (11')

      (00') edge [loop above] node { $\uscore[a]$, .375 } (00')
      (00') edge [bend left=10] node { $\uscore[a]$, .625 } (10')
      (00') edge [loop left] node [above=1,rotate=90] { $\uscore[b]$, .375 } (00')
      (00') edge [bend left=10] node [left=3,above=1,rotate=270] { $\uscore[b]$, .625 } (01')

      (10') edge [loop above] node { $\uscore[a]$, .625 } (00')
      (10') edge [bend left=10] node { $\uscore[a]$, .375 } (00')
      (10') edge [loop right] node [above=1,rotate=270] { $\uscore[b]$, .375 } (10')
      (10') edge [bend left=10] node [above=1,rotate=270] { $\uscore[b]$, .625 } (11')

      (11') edge [loop below] node { $\uscore[a]$, .625 } (11')
      (11') edge [bend left=10] node { $\uscore[a]$, .375 } (01')
      (11') edge [loop right] node [above=1,rotate=270] { $\uscore[b]$, .625 } (11')
      (11') edge [bend left=10] node [left=-3,above=1,rotate=90] { $\uscore[b]$, .375 } (10')

      (01') edge [loop below] node { $\uscore[a]$, .375 } (01')
      (01') edge [bend left=10] node { $\uscore[a]$, .675 } (11')
      (01') edge [loop left] node [above=1,rotate=90] { $\uscore[b]$, .625 } (01')
      (01') edge [bend left=10] node [left=-3,above=1,rotate=90] { $\uscore[b]$, .375 } (00')

      ;
      \end{tikzpicture}
      \caption{Adjacent Inputs}
      \label{figure:adjacent-inputs}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
      \centering
      \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (0) at (-5,  0) { $0$ };
      \node[node] (1) at ( 0,  0) { $1$ };
      \node[node] (2) at ( 5,  0) { $2$ };

      \node[node] (00) at (-4.5, 1.5) { $00$ };
      \node[node] (01) at (-1.5, 2) { $01$ };
      \node[node] (10) at ( 1.5, 2) { $10$ };
      \node[node] (11) at ( 4.5, 1.5) { $11$ };

      \node[node] (00') at (-4.5, -1.5) { $\underline{00}$ };
      \node[node] (01') at (-1.5, -2) { $\underline{01}$ };
      \node[node] (10') at ( 1.5, -2) { $\underline{10}$ };
      \node[node] (11') at ( 4.5, -1.5) { $\underline{11}$ };

      \path
      (00) edge node [left] { $2/3$ } (0)
      (00) edge node [left=50,above] { $1/6$ } (1)
      (00) edge node [left=100,above=15] { $1/6$ } (2)

      (01) edge node [right=30,above=20] { $1/3$ } (0)
      (01) edge node [left=20,above=5] { $1/3$ } (1)
      (01) edge node [left=70,above=20] { $1/3$ } (2)

      (10) edge node [right=70,above=20] { $1/3$ } (0)
      (10) edge node [right=20,above=5] { $1/3$ } (1)
      (10) edge node [left=30,above=20] { $1/3$ } (2)

      (11) edge node [right=100,above=15] { $1/6$ } (0)
      (11) edge node [right=50,above] { $1/6$ } (1)
      (11) edge node [right] { $2/3$ } (2)

      (00') edge node [left] { $2/3$ } (0)
      (00') edge node [left=50,below] { $1/6$ } (1)
      (00') edge node [left=100,below=15] { $1/6$ } (2)

      (01') edge node [right=30,below=20] { $1/3$ } (0)
      (01') edge node [left=20,below=5] { $1/3$ } (1)
      (01') edge node [left=70,below=20] { $1/3$ } (2)

      (10') edge node [right=70,below=20] { $1/3$ } (0)
      (10') edge node [right=20,below=5] { $1/3$ } (1)
      (10') edge node [left=30,below=20] { $1/3$ } (2)

      (11') edge node [right=100,below=15] { $1/6$ } (0)
      (11') edge node [right=50,below] { $1/6$ } (1)
      (11') edge node [right] { $2/3$ } (2)
      ;

      \end{tikzpicture}
      \caption{Randomized Frequency}
      \label{figure:randomized-frequency}
    \end{subfigure}
  \caption{Frequency Estimator}
  \label{figure:frequency-estimator}
\end{figure}

\todo{verify this is $1$-DP}