### 2018-01-28 preliminaries.tex
use standard differential privacy definition e^{\epsilon}

Let $\bbfZ$ and $\bbfZ^{\geq 0}$ be the sets of integers and
non-negative integers respectively. For $n \geq 0$, $\underline{n} =
\{ 0, 1, \ldots, n \} \subseteq \bbfZ^{\geq 0}$.

We briefly review the definitions of differential privacy, Markov
chains, and Markov decision processes. For differential privacy, we
follow the non-standard \todo{why?} definition in~\cite{GRS:09:UUPM,GRS:12:UUPM}. Our
definitions of Markov chains and Markov decision processes are adopted
from~\cite{BK:08:PMC}.

\subsection{Differential Privacy}

We denote the data universe by $\mathcal{X}$; $x \in \mathcal{X}^n$ is
a database with $n$ rows from the data universe. Two databases $x$ and
$x'$ are \emph{neighbors} (denoted by $d(x, x') \leq 1$) if they are
identical except at most one row. A \emph{query} $f$ is a function
from $\mathcal{X}^n$ to its range $R$. The \emph{sensitivity} of the
query $f$ (written $\Delta (f)$) is $\max_{d(x, x') \leq 1} | f (x) -
f (x') |$. For instance, a \emph{counting} query counts the number
of rows with certain attributes (say, female). The sensitivity of a
counting query is $1$ since any neighbor can change the count by at
most one. We only consider queries with finite  ranges for simplicity.
A \emph{data analysis mechanism} (or
\emph{mechanism} for brevity) $M_f$ for a query $f$
is a randomized algorithm with inputs in $\mathcal{X}^n$ and outputs
in $\tilde{R}$.
A mechanism may not have the same output range as its query, that is,
$\tilde{R} \neq R$ in general.
A mechanism $M_f$ for $f$ is \emph{oblivious} if
$\Pr[M_f(x) = \tilde{r}] = \Pr[M_f(x') = \tilde{r}]$ for every
$r \in \tilde{R}$ when $f (x) = f (x')$. In words, outputs of an
oblivious mechanism depend on the query result $f (x)$ but not on the
input database $x$. The order of rows in a database, for instance, is
irrelevant to oblivious mechanisms. A database $x$ is
\emph{$(\epsilon, \delta)$-close} to $x'$ in a mechanism
$M_f$ if for every $\tilde{r} \in \tilde{R}$,
\[
\Pr[M_f (x) = \tilde{r}] \leq e^{\epsilon} \Pr[M_f (x') =
\tilde{r}] + \delta.
%\footnote{The standard definition requires
%$\Pr[M_f (x) = \tilde{r}] \leq e^\epsilon \Pr[M_f (x') =
%\tilde{r}] + \delta$~\cite{DMNS:06:CNSPD,D:06:DP}.
%Since $\epsilon = e^{\ln \epsilon}$, our $(\epsilon, \delta)$-closedness is
%equivalent to the standard $(\ln \epsilon,
%\delta)$-closedness~\cite{GRS:09:UUPM,GRS:12:UUPM}.}
\]
A mechanism $M_f$ is \emph{$(\epsilon, \delta)$-differentially
  private} % $(\epsilon, \delta \approx 0)$
if for every $x, x' \in \mathcal{X}^n$ with $d(x, x') \leq 1$,
$x$ is $(\epsilon, \delta)$-close to $x'$ in $M_f$.

The parameters $\epsilon$ and $\delta$ quantify probabilistically
similar behaviors;
the smaller they are, the behaviors are more similar.
Informally, a differentially private mechamism has probabilistically
similar behabiors on neighboring databases. It will have similar
output distributions when a row is added, deleted, or replaced by
another in a given
database. Consider, for instance, a database contains the row
corresponding to an individual. A differentially private mechanism
will have a similar output distribution when the individual is not in
the database. Regardless of the presence of the individual, such
mechanisms will behave similarly. Privacy of the individual is thus
preserved by differentially private mechanisms. \todo{not clear}

\subsection{Markov Chains and Markov Decision Processes}

Let $\AP$ be the set of \emph{atomic propositions}.
A \emph{(finite) discrete-time Markov chain} $K = (S, \wp, L)$ consists
of a non-empty finite set $S$ of \emph{states}, a \emph{transition
  probability function} $\wp : S \times S \rightarrow [0, 1]$ with
$\sum_{t \in S} \wp(s, t) = 1$ for every $s \in S$, and
a \emph{labeling function} $L : S \rightarrow 2^{\AP}$. A \emph{path}
in $K$ is an infinite sequence $\pi = \pi_0 \pi_1 \cdots \pi_n \cdots$
of states with $\wp (\pi_i, \pi_{i+1}) > 0$ for all $i \geq 0$. We write
$\pi[j]$ for the suffix $\pi_j \pi_{j+1} \cdots$. % Thus $\pi[0] = \pi$.

Let $\Act$ denote a finite set of \emph{actions}.
A \emph{(finite) Markov decision process} $M = (S, \Act, \wp, L)$ consists
of a \emph{transition probability function} $\wp : S \times \Act
\times S \rightarrow [0, 1]$ with $\sum_{t \in S} \wp(s, \alpha, t)
= 1$ for every $s \in S$ and $\alpha \in \Act$ and $S$,
$L$ as for Markov chains. Note the model we consider is \emph{reactive}\footnote{In the literature, such model is referred to as \emph{probabilistic automata}\cite{}. In the model checking context, the notion MDPs is more popular.} as $\Act(s) = \Act$
for every $s \in S$.
A \emph{path} $\pi$ in $M$ is an infinite sequence $\pi_0 \alpha_1
\pi_1 \cdots \pi_n \alpha_{n+1} \cdots$ with
$\wp(\pi_i, \alpha_{i+1}, \pi_{i+1}) > 0$ for all $i \geq 0$.
Similarly, we write $\pi[j]$ for the suffix $\pi_j \alpha_{j+1}
\pi_{j+1} \cdots$ of $\pi$.
\hide{
For any measurable set $\Pi$ of paths, we
write $\Pr[\Pi]$ for the \emph{probability measure} of $\Pi$.
}

Let $M = (S, \Act, \wp, L)$ be a Markov decision process. A
\emph{scheduler} for $M$ is simply an infinite sequence  $\scheduler{S}\in \Act^\omega$.
A path $\pi =
\pi_0 \alpha_1 \pi_1 \cdots \pi_n \alpha_{n+1} \cdots$ is an
\emph{$\scheduler{S}$-path} if $\alpha_{i+1} = \scheduler{S}(\pi_0
\pi_1 \cdots \pi_i)$ for all $i \geq 0$.
Note that a Markov decision process with a
scheduler $\scheduler{S}$ induces a Markov chain $M_{\scheduler{S}} =
(S^+, \wp_{\scheduler{S}}, L')$ where $L' (\sigma s) = L (s)$,
$\wp_{\scheduler{S}} (\sigma s, \sigma s t) = \wp (s,
\scheduler{S}(\sigma s), t)$ for $\sigma \in S^*$ and $s, t \in S$.
\hide{
Subsequently, we write
$\Pr_{\scheduler{S}} (\Pi)$ for the \emph{probability measure} of the
measurable path set $\Pi$ on $M$ given the scheduler $\scheduler{S}$.
}

\hide{
A \emph{reward function} on an MDP $M = (S, \wp, I, L)$ is a function
$r : S \rightarrow \bbfR$. The \emph{reward} for $r (\pi)$ a finite
path $\pi$ in $M$ is $\sum_{i=0}^n s_i$ where $\pi = s_0 \alpha_1 s_1
\cdots s_n$.
}