### 2018-02-01 main.tex

\documentclass{llncs}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{xspace}
\pagestyle{plain}
\usepackage{algorithm,algpseudocode}
\usepackage{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{wrapfig}
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{arrows,calc}
\usepackage{url}
\usepackage{times}

\algtext*{EndWhile}% Remove "end while" text
\algtext*{EndIf}% Remove "end if" text
\algtext*{EndFor}% Remove "end if" text

\input{macros}

\newif\ifpublic
%\publictrue

\title{Model Checking Differentially Private Properties}
\ifpublic
\author{
Depeng Liu\inst{1}
\and
Bow-Yaw Wang\inst{2}
\and
Lijun Zhang\inst{1}}
\institute{
Chinese Academy of Sciences
\and
Academia Sinica
}
\else
\author{Author}
\institute{Institute}
\fi

\begin{document}

\maketitle

\begin{abstract}
  We introduce the branching time temporal logic $\dpCTLstar$ for
  specifying differential privacy. Several differentially private
  mechanisms are formalized as Markov chains or Markov decision
  processes. Using our formal models, subtle privacy conditions
  are specified by $\dpCTLstar$. In order to verify privacy properties
  automatically, model checking problems are investigated.  We
  give a model checking algorithm for Markov chains. Model checking
  $\dpCTLstar$ properties on Markov decision processes however is
  shown to be undecidable.
\end{abstract}

\section{Introduction}
\label{section:introduction}
\input{introduction}

\noindent
\paragraph{Organization of the paper.}
Preliminaries are given in Section~\ref{section:preliminaries}.
In Section~\ref{section:examples} we discuss how differentially private mechanisms can be modeled in Markov chains.
The logic $\dpCTLstar$ and model checking algorithms for Markov chains are presented in
Section~\ref{section:dpCTL}.
Section~\ref{section:specifying-properties} describes differential privacy properties using   $\dpCTLstar$.
They are followed by the extension to Markov decision processes and an
application in stream processing (Section~\ref{section:mdp}). Finally,
Section~\ref{section:conclusions} concludes our presentation.

\section{Preliminaries}
\label{section:preliminaries}
\input{preliminaries}

\section{Markov Chains  for Differentially Private Mechanisms}
\label{section:examples}
\input{examples}

\section{The Logic $\dpCTLstar$}
\label{section:dpCTL}
\input{dpCTL}

\subsection{Syntax}
\label{subsection:syntax}
\input{syntax}

\subsection{Semantics}
\label{subsection:semantics}
\input{semantics}

\section{Specifying Properties in $\dpCTLstar$}
\label{section:specifying-properties}
\input{specifications}

\section{Markov Decision Processes for Differentially Private Mechanisms}
\label{section:mdp}
\input{mdp}

%\section{Model Checking Algorithms}
%\label{section:model-checking-algorithms}
%\input{algorithms}

\subsection{MDPs for Stream Processing}
\label{section:applications}
\input{applications}

\section{Conclusions}
\label{section:conclusions}
We have introduced $\dpCTLstar$ to reason about properties in differential privacy, and have investigated the model checking problem. For Markov chains, the model checking problem has the same complexity as for $\PCTLstar$. The general MDP model checking problem is undecidable. We have discussed that some special cases are decidable. An interesting work is to identify more
decidable subclass. As an example, consider the formula $\bigwedge_{k \in \bbfZ^{\geq 0}}
\dpriv{\epsilon}{\delta} (\X^k  00)$ which was used in specifying the pan privacy. For the case $\epsilon=\delta=0$, it reduces to a language equivalence problem for probabilistic automata. It is interesting to characterize other cases as well.
 Another interesting line of further works is to consider continuous perturbation (such as Laplace distribution used in \cite{DR:14:AFDP}) -- we would need Markov models with continuous state space for the modelling.

\bibliographystyle{splncs03}
\bibliography{refs}

\newpage
\appendix

\section{Proof of Theorem~\ref{theorem:mdp-model-checking}}
\input{proof}

\end{document}