### 2018-04-26 above-threshold.tex
remove streaming and model above threshold to mdp

\subsection{Above Threshold Mechanism}
\label{subsec:threshold}

\begin{algorithm}[tbh]
	\begin{algorithmic}[1]
		\Procedure{AboveThreshold}{$d$, $\{ f_1, f_2, \ldots
                  \}$, $t$}
		\Match{$t$}
		\Comment{obtain $t'$ by $\frac{1}{4}$-geometric mechanism}
		\lCase{$0$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{4}{5}$,$\frac{3}{20}$,
			$\frac{3}{80}$,$\frac{3}{320}$,
			$\frac{3}{1280}$,$\frac{1}{1280}$ respectively}
		\lCase{$1$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{1}{5}$,$\frac{3}{5}$,
			$\frac{3}{20}$,$\frac{3}{80}$,
			$\frac{3}{320}$,$\frac{1}{320}$ respectively}
		\lCase{$2$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{1}{20}$,$\frac{3}{20}$,
			$\frac{3}{5}$,$\frac{3}{20}$,
			$\frac{3}{80}$,$\frac{1}{80}$ respectively}
		\lCase{$3$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{1}{80}$,$\frac{3}{80}$,
			$\frac{3}{20}$,$\frac{3}{5}$,
			$\frac{3}{20}$,$\frac{1}{20}$ respectively}
		\lCase{$4$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{1}{320}$,$\frac{3}{320}$,
			$\frac{3}{80}$,$\frac{3}{20}$,
			$\frac{3}{5}$,$\frac{1}{5}$ respectively}
		\lCase{$5$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{1}{1280}$,$\frac{3}{1280}$,
			$\frac{3}{320}$,$\frac{3}{80}$,
			$\frac{3}{20}$,$\frac{4}{5}$ respectively}
		\EndMatch
		\For{each query $f_i$}
                \State{$r_i \leftarrow f_i (d)$}
		\Match{ $r_i$}
		\Comment{obtain $r_{i}'$ by $\frac{1}{2}$-geometric mechanism}
		\lCase{$0$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{2}{3}$,$\frac{1}{6}$,
			$\frac{1}{12}$,$\frac{1}{24}$,
			$\frac{1}{48}$,$\frac{1}{48}$ respectively}
		\lCase{$1$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{1}{3}$,$\frac{1}{3}$,
			$\frac{1}{6}$,$\frac{1}{12}$,
			$\frac{1}{24}$,$\frac{1}{24}$ respectively}
		\lCase{$2$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{1}{6}$,$\frac{1}{6}$,
			$\frac{1}{3}$,$\frac{1}{6}$,
			$\frac{1}{12}$,$\frac{1}{12}$ respectively}
		\lCase{$3$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{1}{12}$,$\frac{1}{12}$,
			$\frac{1}{6}$,$\frac{1}{3}$,
			$\frac{1}{6}$,$\frac{1}{6}$ respectively}
		\lCase{$4$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{1}{24}$,$\frac{1}{24}$,
			$\frac{1}{12}$,$\frac{1}{6}$,
			$\frac{1}{3}$,$\frac{1}{3}$ respectively}
		\lCase{$5$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
			$\frac{1}{48}$,$\frac{1}{48}$,
			$\frac{1}{24}$,$\frac{1}{12}$,
			$\frac{1}{6}$,$\frac{2}{3}$ respectively}
		\EndMatch
                \State{\textbf{if} $r'_i \geq t'$ \textbf{then} 
                  \textbf{halt} with $a_i = \top$ \textbf{else}
                  $a_i = \bot$}
\hide{
		\If{$r_{i}'>t'$}
		\State{$a_i=\top$}
		\State{\textbf{halt}}
		\Else
		\State{$a_i=\bot$}
		\EndIf
}
		\EndFor
		\EndProcedure

	\end{algorithmic}
	\caption{Input: private database $d$, 
          queries $f_i : \{ 0, 1, \ldots, 5 \} \rightarrow \{ 0, 1,
          \ldots, 5 \}$ with sensitivity $1$, threshold $t \in \{ 0,
          1, \ldots, 5 \}$; Output: $a_1, a_2, \ldots$}
	\label{algorithm:online-model}
\end{algorithm}

\begin{figure}
  \centering
    \resizebox{.8\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (t0r1) at (0, 4.75) { $t_0r_1$ };

      \draw (2.9, 3.95) rectangle (4.9, .05);
      \node at (4.65, 3.25) { $\cdots $};
      \node[node] (t'0r1) at (3.9, 3.25) { $t'_0r_1$ };
      \node at (3.2, 3.25) { $\cdots $};
      \draw (-2.8, 3.95) rectangle (2.8, .05);
      \node[node] (t'1r0) at (2.1, 3.25) { $t'_1r_0$ };
      \node[node] (t'1r1) at (0, 3.25) { $t'_1r_1$ };
      \node[node] (t'1r2) at (-2.1, 3.25) { $t'_1r_2$ };
      \node at (0, 2.1) { $\vdots$ };
      \node at (-2.1, 2.1) { $\vdots$ };
      \draw (-2.9, 3.95) rectangle (-4.9, .05);
      \node at (-3.15, 3.25) { $\cdots$ };
      \node[node] (t'2r1) at (-3.9, 3.25) { $t'_2r_1$ };
      \node at (-4.6, 3.25) { $\cdots$ };

      \node[node] (t'1r'0) at (2.1, .75) { $t'_1r'_0$ };
      \node[node] (t'1r'1) at (0, .75) { $t'_1r'_1$ };
      \node[node] (t'1r'2) at (-2.1, .75) { $t'_1r'_2$ };

      \node[node] (tt0rr0) at (2.1, -4.75) { $\underline{t}_0\underline{r}_0$ };

      \draw (2.9, -3.95) rectangle (4.9, -.05);
      \node[node] (tt'0rr0) at (4.2, -3.25) { $\underline{t}'_0\underline{r}_0$ };
      \node at (3.35, -3.25) { $\cdots$ };
      \draw (.8, -3.95) rectangle (2.8, -.05);
      \node[node] (tt'1rr0) at (2.1, -3.25) { $\underline{t}'_1\underline{r}_0$ };
      \node at (1.25, -3.25) { $\cdots$ };

      \draw (.7, -3.95) rectangle (-4.9, -.05);
      \node[node] (tt'2rr0) at (0, -3.25) { $\underline{t}'_2\underline{r}_0$ };
      \node[node] (tt'2rr1) at (-2.1, -3.25) { $\underline{t}'_2\underline{r}_1$ };
      \node[node] (tt'2rr2) at (-4.2, -3.25) { $\underline{t}'_2\underline{r}_2$ };

      \node at (-4.2, -1.9) { $\vdots$ };

      \node[node] (tt'2rr'0) at (0, -.75) { $\underline{t}'_2\underline{r}'_0$ };
      \node[node] (tt'2rr'1) at (-2.1, -.75) { $\underline{t}'_2\underline{r}'_1$ };
      \node[node] (tt'2rr'2) at (-4.2, -.75) { $\underline{t}'_2\underline{r}'_2$ };

      \path
      (t0r1) edge [bend left=20] node [above] { $\frac{1}{1}$ } (t'0r1)
      (t0r1) edge node [above] { $\frac{1}{1}$ } (t'1r1)
      (t0r1) edge [bend right=20] node [below] { $\frac{1}{1}$ } (t'2r1)

%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'0)
%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'1)
%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'2)

      (t'1r'0) edge [double] node [above] { $0|0, 0|1$ } (t'1r0)
      (t'1r'0) edge [double, bend left=10] node [above] 
              { $1|0, 1|1, 1|2$ } (t'1r1)
      (t'1r'0) edge [double, bend left=20] node [below] 
              { $2|1, 2|2$ } (t'1r2)

      (tt0rr0) edge [bend right=20] node [above] { $\frac{1}{1}$ } (tt'0rr0)
      (tt0rr0) edge node [above] { $\frac{1}{1}$ } (tt'1rr0)
      (tt0rr0) edge [bend left=20] node [below] { $\frac{1}{1}$ } (tt'2rr0)

      (tt'2rr'0) edge [double] node [above] { $0|0, 1|0$ } (tt'2rr0)
      (tt'2rr'0) edge [double] node [above] { $0|1, 1|1, 2|1$ } (tt'2rr1)
      (tt'2rr'0) edge [double] node [above] { $1|2, 2|2$ } (tt'2rr2)

      (tt'2rr'1) edge [double] node [above] { $0|0, 1|0$ } (tt'2rr0)
      (tt'2rr'1) edge [double] node [above] { $0|1, 1|1, 2|1$ } (tt'2rr1)
      (tt'2rr'1) edge [double] node [above] { $1|2, 2|2$ } (tt'2rr2)

      ;
      \end{tikzpicture}
    }
  \caption{Markov Decision Process for Above Threshold}
  \label{figure:mdp-above-threshold}
\end{figure}

In differential privacy, an offline mechanism releases outputs only
once and plays no further role; an online (interactive) mechanism
allows analysts to ask queries adaptively based on previous
responses. The mechanisms we constructed previously are offline
mechanisms. On the other hand, online mechanisms permit further queries. 
Obviously, queries do not change databases themselves. To maintain
privacy, online mechanisms may decide to release
information differently. For instance, they may disable further
queries after a certain output is disclosed.

Below we describe an online mechanism adapted from~\cite{DR:14:AFDP}.
Given a threshold and a series of adaptive queries, we only care for
the queries whose results are above the threshold; queries below the
threshold only disclose minimal information and hence are considered
irrelevant. Since query results above the threshold convey useful
information, let us assume the mechanism will halt on the first such
query result for simplicity. The question is to design a
privacy-preserving mechanism to have similar behaviors on similar
databases. In~\cite{DR:14:AFDP}, a mechanism is design for continuous
queries by applying the Laplace mechanism. We will develop a mechanism
for bounded and discrete queries using the geometric mechanism.

Consider a threshold $t \in
\{ 0, 1 \ldots, 5\}$, and queries $\{ f_i : f_i $ with sensitivity $\Delta
(f_i) = 1 \}$. In order to protect privacy, our mechanism applies the
truncated $\frac{1}{4}$-geometric mechanism to obtain a noisy
threshold $t'$. For each query, the truncated $\frac{1}{2}$-geometric
mechanism is applied to its result. If the noisy result is not less
than the noisy threshold, our mechanism outputs $\top$ and
halts. Otherwise, it outputs $\bot$ and continues to the next query
(Algorithm~\ref{algorithm:online-model}). The above threshold
mechanism outputs a sequence of the form $\bot^* \top$. On
similar databases, we would like to show that the discrete above
threshold mechanism outputs the same sequence with similar
probabilities.