### 2018-04-26 subsampling.tex
add subsampling algorithm

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{Subsampling}{$d$, $f$}
    \Require{$d \in \{ R, B \}^3$, $f : \{ R, B \}^3 \rightarrow \{ R,
      B \}$}
    \State{Subsample $m$ data sets $\hat{d}_1, \hat{d}_2, \ldots,
      \hat{d}_m$ from $d$ where each row of $d$ is chosen with
      probability $q$}
    \State{$z \leftarrow \{ f (\hat{d}_1), f (\hat{d}_2), \ldots, f
      (\hat{d}_m) \}$}
    \State{$\mathit{count}_R, \mathit{count}_B \leftarrow 
            | \{ i : f (\hat{d}_i) = R \} |,
            | \{ i : f (\hat{d}_i) = B \} |$}
    \State{$r \leftarrow | \mathit{count}_R - \mathit{count}_B | /
      (4mq) - 1$}
    \State{$\hat{r} \leftarrow r + \mathit{Lap}(\frac{1}{\epsilon})$}
    \If{$\hat{r} > \ln (1/\delta)/\epsilon$}
      \State{\textbf{if} $\mathit{count}_R \geq \mathit{count}_B$
             \textbf{then} \Return { $R$ } \textbf{else} \Return {$B$} } 
    \Else
      \State{\Return {$\bot$}}
    \EndIf
  \EndFunction
  \end{algorithmic}
  \caption{Subsampling}
  \label{algorithm:subsampling}
\end{algorithm}