### 2018-05-02 subsampling.tex
add description of subsampling

The sensitivity of queries is required to apply the (truncated)
$\alpha$-geometric mechanism. Recall that the sensitivity of a query
is the maximal difference of query results on any two neighbors. There are
two problems for such mechanisms depending on query sensitivity. First, 
sensitivity of queries can be hard to compute or estimate. Second,
the sensitivity over arbitrary neighbors can be too conservative for
the actual dataset in use. One therefore would like to have
mechanisms independent of query sensitivity.

Subsampling is a technique to design such mechanisms. Concretely, let
us consider $\mathcal{X} = \{ R, B \}$ (for red and blue teams) and a
dataset $d \in \mathcal{X}^n$. Suppose we would like to ask which team
is the majority in the dataset while respecting individual
privacy. This can be achieved as follows. The mechanism first samples $m$
sub-datasets $\hat{d}_1, \hat{d}_2, \ldots, \hat{d}_m$ from
$d$. It then computes $m$ query results from the $m$
sub-datasets. Finally, the mechanism reports a perturbed result in a
privacy respecting way
(Algorithm~\ref{algorithm:subsampling-majority}).

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{SubsamplingMajority}{$d$, $f$}
    \Require{$d \in \{ R, B \}^n$, $f : \{ R, B \}^* \rightarrow \{ R,
      B \}$}
    \State{$q, m \leftarrow 
            \frac{\epsilon}{64\ln (1/\delta)},
            \frac{\log (n/\delta)}{q^2}$}
    \State{Subsample $m$ data sets $\hat{d}_1, \hat{d}_2, \ldots,
      \hat{d}_m$ from $d$ where each row of $d$ is chosen with
      probability $q$}
    \State{$\mathit{count}_R, \mathit{count}_B \leftarrow 
            | \{ i : f (\hat{d}_i) = R \} |,
            | \{ i : f (\hat{d}_i) = B \} |$}
    \State{$r \leftarrow | \mathit{count}_R - \mathit{count}_B | /
      (4mq) - 1$}
    \If{${r} + \mathit{Lap}(\frac{1}{\epsilon}) > \ln (1/\delta)/\epsilon$}
      \State{\textbf{if} $\mathit{count}_R \geq \mathit{count}_B$
             \textbf{then} \Return { $R$ } \textbf{else} \Return {$B$} } 
    \Else
      \State{\Return {$\bot$}}
    \EndIf
  \EndFunction
  \end{algorithmic}
  \caption{Subsampling Majority}
  \label{algorithm:subsampling-majority}
\end{algorithm}

\begin{figure}
  \centering
    \resizebox{.9\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]

      \node[node] (rrr) at (-5,  2.25) { $rrr$ };
%      \node at (-4.4, 2) { $\vdots$ };
      \node[node] (rrb) at (-5,  0.75) { $rrb$ };
%      \node at (-4.4, .75) { $\vdots$ };
      \node[node] (rbb) at (-5, -0.75) { $rbb$ };
%      \node at (-4.4, -.5) { $\vdots$ };
      \node[node] (bbb) at (-5, -2.25) { $bbb$ };
%      \node at (-4.4, -1.75) { $\vdots$ };

      \node[node] (rmb0) at (-2.5,  2.25) { $m,0$ };
      \node at (-2.5, 0) { $\vdots$ };
      \node[node] (r0bm) at (-2.5, -2.25) { $0,m$ };

      \node[node] (Rm) at (0,  1.5) { $r_m$ };
      \node at (0, 0) { $\vdots$ };
      \node[node] (RM) at (0, -1.5) { $r_M$ };

      \node[node] (top) at (2.5,  0.75) { $\top$ };
      \node[node] (bot) at (2.5, -0.75) { $\bot$ };

      \node[node] (outR) at (5,  1.25) { $R$ };
      \node[node] (outB) at (5, .25) { $B$ };

      \path
      (rrr) edge node [above] { } (rmb0)
      (rrr) edge node [above] { } (r0bm)
      (rrb) edge node [above] { } (rmb0)
      (rrb) edge node [above] { } (r0bm)
      (rbb) edge node [above] { } (rmb0)
      (rbb) edge node [above] { } (r0bm)
      (bbb) edge node [above] { } (rmb0)
      (bbb) edge node [above] { } (r0bm)

      (rmb0) edge node [above] {} (RM)
      (r0bm) edge node [above] {} (RM)

      (Rm) edge node [above] {} (top)
      (Rm) edge node [above] {} (bot)
      (RM) edge node [above] {} (top)
      (RM) edge node [above] {} (bot)

      (top) edge (outR)
      (top) edge (outB)
      ;

      \end{tikzpicture}
    }
  \caption{Markov Chain for Subsampling Majority}
  \label{figure:mc-subsampling}
\end{figure}