### 2018-06-08 dpCTL.tex
polish

The logic $\dpCTLstar$ is designed to specify differentially private
mechanisms. We introduce the differentially private path quantifier
$\dpriv{\epsilon}{\delta}$ and neighborhood relations for neighbors in
$\dpCTLstar$. For any path formula 
$\phi$, a state $s$ in a Markov chain $K$ satisfies
$\dpriv{\epsilon}{\delta} \phi$ if the probability of having paths 
satisfying $\phi$ from $s$ is close to the probabilities of
having paths satisfying $\phi$ from its neighbors.
%We give semantics for both Markov chains and Markov decision processes.