### 2018-07-27 rebuttal-aplas.txt
add reviews and rebuttals

First of all, we would like to thank all reviewers' efforts and
comments.

The main contributions of this work are the formalization of privacy
protection mechanisms as Markov chains or Markov Decision Processes,
and its automated analyses. We choose to formalize interactive
mechanisms by non-determinism in Markov Decision Processes and
illustrate our formalization by the above threshold mechanism
(Algorithm 3) for clarity. In the above threshold mechanism, the
incoming queries f_i's are selected dynamically. Later queries may
depend on the results of previous queries. Hence the interaction is
implicit in the choices of incoming queries. It would be great if we
could give our formalization of more explicit interactions such as
those in streaming data. Due to space limit, we had to present a
simpler mechanism with implicit interactions. We also agree a thorough
investigation on the expressiveness of our formalization will be very
useful. It will definitely be an interesting future work. Our
(un)decidability results are for dpCTL*. As long as a mechanism can be
modeled by Markov chains, dpCTL* properties can be verified
automatically. Model checking dpCTL* over Markov Decision Processes on
the other hand is undecidable. Hence interactive mechanisms in general
cannot be verified with our current formalization.

We are certainly aware of various language-based techniques for
verifying differential privacy. Just like classical
programs/properties, we see both techniques are complementary.
Classical model checking aims to reduce human efforts by focusing on
abstract models with automated verification. It has helped programmers
verify simpler programs and properties reasonably successfully over
decades of research efforts. Our work has the same position. It does
not aim to verify all possible differentially private programs.
Instead, it hopefully will solve simpler problems so that programmers
can have more time for language-based techniques.